---
title: "AidData Vignette"
author: "Felix Haass"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AidData Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

## AidData for R

The [AidData](http://aiddata.org) project makes available a wealth of information
on development finance. At the core of their database lies the project-level information
about development financing. The `aiddata` package for `R` simplifies the 
access to this database directly from the `R` console.

## Installation

`aiddata` is not (yet) published on CRAN. It needs to be installed directly
from GitHub, using `devtools`:

```{r, eval = FALSE}

install.packages("devtools")
library(devtools)
install_github("felixhaass/aiddata", dependencies = TRUE)
library(aiddata)

```


## Overview

`aiddata` is a set of functions that are wrapped around [`httr`](http://cran.r-project.org/web/packages/httr/) functions to access and download data from the [AidData API](http://aiddata.org/use-aiddatas-api). 

The main functions are:

* `get_aid()`: Download project-level information for one or more countries in a given time period.  
* `get_gis()`: Download geographic point information for a subset of aid projects (not all aid projects are geo-coded)  
* `browse_aid()`: Open browser page with detailed project information for a single project.

## Examples

`get_aid()` takes as input an ISO-2 character string of recipient or donor country
and start and end years. It downloads information on all aid projects in the given 
recipient or from the given donor. If both recipient and donor are provided, the resulting 
aid projects represent information on aid flows in the recipient-donor dyad. 

ISO-2 codes for donors and recipient can be obtained with the [`countrycode`](http://cran.r-project.org/web/packages/countrycode/index.html) package.

```{r, eval = FALSE}

result <- get_aid(rec = "AO", start = 1990, end = 1995)

```



